Relationships in the extremes and their influence on competition and coexistence

### 2022-01-10 paper/Paper.Rnw
formatting change for figure numbers in the sup mat
edits to G4 methods content and new figure caption, not yet done though

%***DAN: Methods paras here for G4, if we accomplish G4
To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we consider the coexistence of diatom 
species established by chemostat batch experiments \citep{Descamps_2005}.
\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
resource model as follows:
put model here
Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C. pseudostelligera} 
and \emph{F. crotonensis} occurs with oscillating temperature fluctuations, 
$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, defined by parameters $\theta_0 = 18, a = 6, P = 60$d. 
\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, 
and here, we consider varying parameters of temperature fluctuations and further decompose the 
coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
environmental fluctuation is temperature fluctuation. We first considered the scenario where 
\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
the coexistence decomposition for \emph{F. crotonensis}. 
%put reverse in sup mat?
%do we need to explain how we separated the asymmetries using the sims??
Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
$P$, a total of three plots. For each plot, we set two of the parameters constant at the 
original parameters which coexistence was experimentally found in \cite{Descamps_2005}, 
and let the remaining temperature parameter vary. 
We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully 
large for the lottery model, compared to 
the contributions of other mechanisms. 
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).

To further explore the parameter space, we generated two dimensional surface plots which reveal 
the subsets of parameter space where coexistence is possible with ranging degrees of ATA 
contributions and where ATA rescue occurs. We evaluated coexistence under two varying 
parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total 
IGR, with a color bar and contour lines. Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and 
higher color intensity indicate greater contribution of ATAs to the IGR and positive cases 
where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
(\ref{eq:full_decomp})], and the difference of these quantities. 
For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
``trait-difference space'' corresponding to ATA rescue 
($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
$\text{IGR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[EC]}$ is positive 
correspond to ATA exclusion (shaded violet). Top panels also include domains
of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
probably of minimal practical importance. 
Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
so are not displayed. The dashed red line on panel L corresponds to the parameters
used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
\end{figure}
%***DAN: Dan has edited this caption

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{The contribution of ATAs to diatom coexistence is meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
two of the three parameters defining temperature oscillation, amplitude, period, and mean, 
is fixed at the original coexistence values found by \cite{Descamps_2005} ($a=6, P=60, \theta_0=18$). The IGR and components of the decomposition
are plotted against one varying parameter. The original value of the varying parameter and coexistence quantities at that value is indicated by the vertical dashed magenta lines. The contribution of ATAs (red line)
has magnitude comparable to other components and, in these parameters, is always positive, indicating that ATAs can contribute meaningfully
to diatom coexistence. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.}  \label{fig:fig5}
\end{figure}

>>>>

%***DAN: Methods paras here for G4, if we accomplish G4
To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we consider the coexistence of diatom 
species established by chemostat batch experiments \citep{Descamps_2005}.
\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
resource model as follows:
put model here
%***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C. pseudostelligera} 
and \emph{F. crotonensis} occurs with oscillating temperature fluctuations, 
$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, 
%***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
defined by parameters $\theta_0 = 18, a = 6, P = 60$d. 
\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, 
and here, we consider varying parameters of temperature fluctuations and further decompose the 
coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
environmental fluctuation is temperature fluctuation. We first considered the scenario where 
\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
the coexistence decomposition for \emph{F. crotonensis}. 
%put reverse in sup mat?
%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to just consider 1 invading 2 for 
%the lottery model, but cannot really do that here, right?
%do we need to explain how we separated the asymmetries using the sims??
%***DAN: I am not sure what you mean by this question, Jasmin
As for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
$P$, a total of three plots. For each plot, we set two of the parameters constant at the 
original parameters which coexistence was experimentally found in \cite{Descamps_2005}, 
and let the remaining temperature parameter vary. 
We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully 
large for the lottery model, compared to 
the contributions of other mechanisms. 
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).

To further explore model parameter space, we generated two dimensional surface plots which reveal 
some subsets of parameter space where coexistence is possible with ranging degrees of ATA 
contributions and where ATA rescue occurs. We evaluated coexistence under two varying 
parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total 
IGR, with a color bar and contour lines. Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and 
higher color intensity indicate greater contribution of ATAs to the IGR and positive cases 
where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
(\ref{eq:full_decomp})], and the difference of these quantities. 
For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
``trait-difference space'' corresponding to ATA rescue 
($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
$\text{IGR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[EC]}$ is positive 
correspond to ATA exclusion (shaded violet). Top panels also include domains
of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
probably of minimal practical importance. 
Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
so are not displayed. The dashed red line on panel L corresponds to the parameters
used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
\end{figure}

%***DAN: Jasmine, some changes are needed to the figure itself. Please make the text in the legend bigger.
%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
%(that's what it says in the legend),
%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
%"A".
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} 
is meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
%invading which, and, at the end of the 
%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
For each panel, 
two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
and mean, $\theta_0$) are fixed at the original coexistence values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, \theta_0=18^\circ$C). 
%***DAN: I added units for a  and \theta_0 here, please check these are correct 
The IGR and components of the decomposition
are then plotted against the other parameter. The original value of the varying parameter and coexistence quantities at that value is indicated by the vertical dashed magenta lines. The contribution of ATAs (red line)
has magnitude comparable to other components and, in these parameters, is always positive, indicating that ATAs can contribute meaningfully
to diatom coexistence. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.
%***DAN: Don't forget to fill in the standard error stuff
}  \label{fig:fig5}
\end{figure}
